---
title: "secuTrial R package vignette"
author: "Patrick R. Wright"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This R package provides functions for handling data from the clinical data management system secuTrial. The most important components are
related to loading data exports from secuTrial into R. For this vignette and also to test your installation of the R secuTrial package
some exaple export data is delivered with the package. The data (*calcium*) was retrieved from the **lava** package and prepared for the 
import into a secuTrial eCRF. It was then exported and added to the secuTrial R package.

## Load the secuTrial library

```{r}
library(secuTrial)
```

## Loading data from a rectangular export

```{r}
load.tables(data.dir=system.file("extdata", "s_export_rt-CSV-xls_DEM00_20181016-151332.zip", package = "secuTrial"), 
            is.rt = T, 
            decode.rt.visitlabels = T)
```

The above command loads the rectangular export and writes it into the variable *rtdata*. 
Furthermore, the visitlabels are automatically translated if *decode.rt.visitlabels* is set to TRUE.

## Loading data from a non-rectangular export
```{r}
load.tables(data.dir=system.file("extdata", "s_export_CSV-xls_DEM00_20180912-125720.zip", package = "secuTrial"), tables=c("bmd.xls"))
```

The above command reads export tables into R variables. The central table in this case is *bmd* which contains the form data of the bone mineral density
dataset. The *tables* argument has been set to prevent loading of all tables.

## Column moving

```{r}
bmd_subset <- bmd[,c("pat.id","mnpvispdt","age","grouping","bmd")]
names(bmd_subset)
# position "grouping" and "bmd" behind "mnpvispdt"
bmd_subset_moved_1 <- move.column.after(df=bmd_subset,col.name=c("grouping","bmd"),"mnpvispdt")
names(bmd_subset_moved_1)
# position "age" back to position 3 with "move.column.to.pos"
bmd_subset_moved_2 <- move.column.to.pos(df=bmd_subset_moved_1,col.idx=5,new.col.idx=3)
names(bmd_subset_moved_2)

```

## Translating IDs

```{r}
mnppid2mnpaid(1512)
mnpaid2mnppid(104)
```

## Stripping center tags
```{r}
remove.center.tag("Universitätsspital Basel (SWISS-AF)")
remove.center.tag("HUG Genève (SSR)")
```

## Retrieve center from mnppid
```{r}
mnppid2center(1509)
mnppid2center(1509, remove.ctag = 0)
```

## Adding center to tables
```{r}
bmd_with_center_col <- add.center(bmd)
length(names(bmd))
length(names(bmd_with_center_col))
```

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
